---
title: "Determinants of crime in the world"
author: "Joanna Ceglińska, Szymon Groszkiewicz"
date: "6 maja 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Project note

The aim of this project is to reproduce the results of a bachelor thesis study prepared in 2019 and possibly extend in using more advanced methods of data analysis and visualization. 
The topic of the research was investigating the determinants of crime using panel data of 48 countries for the years 2003-2015. The dependent variable was the homicide rate and independent variables include the Gini index, GDP per capita, unemployment rate, variables responsible for measuring the education level in a country and a regressor which presents the effectiveness of judicial system (Unsentenced detainees as a proportion of overall prison population).

The summary of variables is presented below:

```{r}
df = read.csv("homicide_data.csv", sep=",", header=TRUE)
summary(df)
```

